Estimating the likelihood of student replication success after an initial failure to replicate

### 2024-03-12 manuscript/manuscript.Rmd

---
title: "Estimating the likelihood of student replication success after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: TODO x a lot
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt

header-includes:
 - \usepackage{tikz}
 - \usetikzlibrary{positioning,chains}
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.00}
 - \renewcommand{\topfraction}{1}
 - \renewcommand{\bottomfraction}{1}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}

---

@@

::: {.abstract data-latex="" lang=en}
<!-- In order to do cumulative science, results need to be replicable. However, an individual direct replication could get a different result than an original ("replication failure") for a number of reasons, which vary in whether they would apply to a potential second replication. Given one failed replication, how likely is a second attempt at replication to succeed?-->

As a pragmatic matter, when a replication fails, scientists have to decide: do they make a second attempt at replication, or move on? 
Given the relatively low rates of replication success reported in large scale replication projects, this choice point is one facing many scientists in psychology who attempt to replicate a study before building on it. Here, we report on 17 re-replications of experiments with failed replicated reported in  Boyce et al. (2023). In 5/17 of these rescue projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes. While a few studies with failed replications were rescued with larger sample sizes and experimental tweaks, our overall results indicate that most failed replications stayed failed. If the goal is to do cumulative science, one may wish to stop after one failed replication.

:::

<!-- Other Title ideas:

@@

Estimating the likelihood of [student] replication success after an initial failure to replicate
--> 
# Introduction

imagine a scenario when this happens (motivating example). Imagine the scenario where you are a new graduate student, and you run a replication of a study in the literature that you are interested in building upon in your research. The replication fails: perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a p-value of .3. Or perhaps the interaction is numerically in the wrong direction, and also the main effects look different. Or perhaps the main effect of interest is null, but you found a significant effect on the interaction that wasn't hypothesized. Whatever the details of the replication failure, you are left with a question: Should you *try again* and run a re-replication, or should you *give up* and pick a different study to replicate and build on?

We focus on scientists’ decisions about what studies to run and how to expend their limited resources in the face of an uncertain literature. Scientists may replicate studies for a variety of reasons, including calibrating their equipment, confirming a result before building on it, and engaging in meta-scientific practices [other reasons to highlight? is there any citation for this?].
We focus on this middle motivation, where a scientist wants to build on a result, and thus wants it to succeed, but does not have strong reasons to believe or disbelieve the finding: it isn't a robust, well-tested effect, nor is it something they have particular suspicions about. However, in an uncertain literature, we might not want to commit resources to building on an effect without first checking that we, with our resources and participant sample, can get the core effect we wish to build on. Thus, scientists may want to re-replicate failed replications if the re-replication is likely to work and is indicative that future experiments in this tradition are likely to be informative.

@@

# Methods
PSYCH 251 is a graduate-level experimental methods class taught by MCF. In previous years, students have conducted replication projects, as reported in @boyce2023. In Fall 2023, students in PSYCH 251 were offered the option to do a "rescue" project where they re-replicated one of the unsuccessful replications from a previous year (students could also opt to do a normal replication instead). We report on the result of `r num_proj` rescue projects that opted to be part of the paper and completed data collection.

@@

## Sample
The experiments that were re-replicated were a non-random sample of studies from @boyce2023. 
We created an initial list of `r nrow(filtered)` rescue-eligible studies that had received a subjective replication success score of 0, .25 or .5 (on a 0-1 scale) in @boyce2023, where the replication had a github repository available (github repositories were used starting in academic year 2015-2016), and where the original experiment had 200 or fewer participants (for feasibility reasons if we needed to increase power). We then contacted the replication project authors for permission to share their report and github repository with a new student and include it as a supplement on a resulting paper. This left `r length(approved)` options that were offered to the students. `r length(chosen)` students chose to do rescue projects; `r length(chosen)-num_proj` students took an incomplete or did not indicate interest in being part of the rescue paper, leaving a final sample of `r num_proj` rescue projects.

@@

## Procedure 
Students conducted their rescue projects over the course of the 10-week class. Once they had chosen a project we gave them access to the original replicators' write-up and repository, which often included the data, experiment code, and analytic code. In many cases, students were also given the contact information of the original replicator (a few original replicators opted not to be contacted by students).

@@

With one exception, samples were collected on Prolific (the rescue of @yeshurun2003 ran in-person on the Stanford student subject pool). We tried to power studies adequately (with a target of 2.5x original following @simonsohn2015), but due to cost constraints, not all studies were powered at this level. The rescues had on average `r round(m,2)` times the original sample post-exclusion (median: `r round(mid,2)`, IQR: `r round(low,2)` - `r round(high,2)`, minimum: `r round(lowest, 2)`, maximum: `r round(highest,2)`, see Table \@ref(tab:tab-size) for all sample sizes). 
Across the `r num_proj-1` Prolific studies, we spent $`r pluck(cost,1,1) |> round()`, for an average of $`r (pluck(cost,1,1)/(num_proj-1)) |> round()` per project.

## Pre-registration
Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects. Each project was also individually pre-registered by the student conducting it. The overall analysis is at LINK, individual pre-registrations are linked from LINK.

@@

## Coding of results
We followed @boyce2023 in what study properties we measured and what measures of replication success we used.

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP. Disagreements were resolved through discussion. As a compliment to the overall subjective rating of success, we followed @boyce2023 in also doing a statistical comparison on one key measure of interest for each study. In order to statistically compare the key measures, we needed effect sizes reported in the same way within each original-replication cluster. When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We also recorded the same set of potential correlates that were used in @boyce2023 for original, replication, and rescue (these were already rated for original and replication). These potential correlates included features of the original study including the subfield of the study (cognitive v social v other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, and how many items each participant saw, and whether there were open materials and open data.

@@

## Overall replication rate 
<!--* We will report the distribution of subjective replication success in our rescue sample

* If there is a mixture of projects that succeed and fail to replicate the original results, we will qualitatively describe differences that may have played a role. -->
All rescue projects were rated holistically for how well they replicated the original results. We thought about replication in terms of how confident one would be to build on the line of work given the replication results, rather than focusing on any singular numeric result or significance cut-off.  All projects were rated both by the instructor (MCF) and by one of the TAs (VB or BP); the interrater reliability was `r irr |> round(3)`.

@@

Another successful replication was that of @ngo2019. Here, the original study found a large effect, and so the first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people, and then failed to find the effect. The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one). There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separately potential reasons that slight changes to experiments, or switches to online, could result in noisier samples (and thus smaller effect sizes). Thus, replications with smaller samples than the original (even if powered to the original effect size), may not be that diagnostic, and could potentially benefit from a re-replication.

@@

(TODO other successes or failures we want to discuss? )

@@

Small replication samples relative to original and rescue could be due both to a) aiming for a small replication sample due to aiming for power for a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample. Since relative sizes of the studies may play a role in replication success and how probitive replications are, we show the sample sizes in Table \@ref(tab:tab-size).

@@

Overall, we do not have a clear picture of why certain studies replicated in the rescue sample and others did not, aside from a couple cases where fixing sample size issues may have helped.

@@

As a complement to the holistic, qualitative replication ratings used above, we also statistically compared the effect sizes of the rescue, original replication, and original study on one key measure per study. We followed @boyce2023 in the determination of a key measure for each study.  When we were aware of additional direct replications (either from other class projects, or external replications in the literature), we also consider the effect sizes obtained in these additional replications.

We standardized all effect sizes into standardized mean difference (SMD) units. One potential issue with comparisons using SMD is that noisier measures will have smaller standardized effect sizes even if the effect on the original scale is the same. In general, the replication and rescue effect sizes were smaller than the original effect sizes, and in a couple cases the effects were in the opposite direction (Figure \@ref(fig:smd)).

@@

Our intuitions of replication and intuitions of whether an effect provides support for a hypothesis (including heuristic cutoffs like p<.05) don't always align with measures of statistical consistency [@patil2016]. For instance, two studies may both find that condition 1 results in a significantly higher outcome measure than condition 2, but the effect magnitudes may be sufficiently different that they are statistically unlikely to have come from the same population. On the other side, one study may find a statistically significant, but imprecisely estimated effect in a small sample, and second study may find a near-zero (null) effect, but the effect estimates from the two studies may be statistically compatible, despite one supporting a hypothesized difference and the other not.

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that two effects come from the same distribution. As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found). Because we have few replication for any study, we impute the heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

@@

As secondary measures, we also calculated the $p$-original values for a) the original and the rescue, b) the original and non-rescue replications, and c) between the rescue and other replications. 
For a) the rescue versus the original, median value of $p$-original was `r rescue_summ$median |> round(2)` [IQR: `r rescue_summ$low |> round(2)` - `r rescue_summ$high|> round(2)`], and  `r round(rescue_summ$less_than*100)`% of the $p$-original values were less than .05.
For b) the non-rescue replications versus the original, median value of $p$-original was `r rep_summ$median |> round(2)` [IQR: `r rep_summ$low |> round(2)` - `r rep_summ$high |> round(2)`], and  `r round(rep_summ$less_than*100)`% of the $p$-original values were less than .05. For c) the rescue versus the other replications, median value of $p$-original was `r rescue_rep_summ$median |> round(2)` [IQR: `r rescue_rep_summ$low |> round(2)` - `r rescue_rep_summ$high |> round(2)`], and  `r round(rescue_rep_summ$less_than*100)`% of the $p$-original values were less than .05.

@@

<!--* We will use p-original to evaluate how consistent the original effect size is with the totality of replications. We expect there to be a small number of replications, so we will impute the heterogeneity value as in https://osf.io/preprints/psyarxiv/dpyn6/. --> 
Overall, allowing for a heterogeneity level of $\tau$=.21 SMD, a number of rescues and replications are not statistically consistent with the original effects. The pattern of inconsistency does not align with which studies were rated as having replicated. In all cases, the point estimate of the re-replication is smaller than (or in the opposite direction) of the original effect on the key measure of interest.

@@

# Discussion
We presented the results of `r num_proj` new replications that attempted to "rescue" previous failed replications reported in @boyce2023 by identifying possible causes of non-replication and ameliorating them in the re-replication. `r nrow(success)` of these re-replications (`r round(nrow(success)/num_proj*100)`%) mostly or fully replicated.

We don't have qualitative or quantitative explanations for why the specific studies did or did not replicate. In a couple cases, increasing sample size and fixing internal validity issues in the replication seems to have led to a successful rescue (although we can't establish causality even in these cases). However, there were other cases where the original replications had issues with a small sample or deviations in the implementation, and the re-replication addressed these issues but still failed to replicate the original results. We can't predict what replication failures are likely to resolve given another, more thoughtful try, beyond that suggestion that glaring problems and low samples may sometimes (but not always) be resolvable.

Across all the re-replications, regardless of whether the pattern of effects replicated or not, the effect sizes were smaller than the original effect, often by a substantial amount. In a substantial minority of the replications, the replication effect sizes (taken as a whole) were statistically inconsistent with the original effect, even allowing for some heterogeneity between studies. These diminished and inconsistent effects suggest that even if a re-replication "works", it may be difficult to build upon as follow-up studies will need large samples to detect the small effects.

The reported rescue projects are a small sample of replications, and the effects explored are non-random, as they have been doubly selected for by student interest. However, this selection bias is likely to correlate with how graduate students choose what topics to work on and what studies to build on. Nonetheless, with a small sample all our estimates are highly uncertain.

@@

We opened this paper with a question about what an early-career psychology researcher should do given a failed replication: should they try again or move on? From our sample of testing the "try again" approach, it seems that the odds of a re-replication working are low (consistent with @ebersole2020), so especially if there isn't a super clear failure mode to point to, another try probably won't rescue the original results.

# Acknowledgements {-}

# Author Contributions {-}

# References {-}

<!-- Use this magic to place references here. -->
<div id="refs"></div>

>>>>

---
title: "Estimating the likelihood of student replication success after an initial failure to replicate"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: TODO x a lot
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251rescue.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: apa7.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt
header-includes:
 - \usepackage{tikz}
 - \usetikzlibrary{positioning,chains}
 - \usepackage{setspace}\singlespacing
 - \renewcommand{\textfraction}{0.00}
 - \renewcommand{\topfraction}{1}
 - \renewcommand{\bottomfraction}{1}
 - \renewcommand{\floatpagefraction}{1}
 - \setcounter{topnumber}{3}
 - \setcounter{bottomnumber}{3}
 - \setcounter{totalnumber}{4}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

::: {.abstract data-latex="" lang="en"}
<!-- In order to do cumulative science, results need to be replicable. However, an individual direct replication could get a different result than an original ("replication failure") for a number of reasons, which vary in whether they would apply to a potential second replication. Given one failed replication, how likely is a second attempt at replication to succeed?-->

As a pragmatic matter, when a replication fails, scientists have to decide: do they make a second attempt at replication, or move on?
Given the relatively low rates of replication success reported in large scale replication projects, this choice point is one facing many scientists in psychology who attempt to replicate a study before building on it.
Here, we report on 17 re-replications of experiments with failed replicated reported in Boyce et al. (2023).
In 5/17 of these rescue projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes.
While a few studies with failed replications were rescued with larger sample sizes and experimental tweaks, our overall results indicate that most failed replications stayed failed.
If the goal is to do cumulative science, one may wish to stop after one failed replication.
:::

```{=html}
<!-- Other Title ideas:

@@

Estimating the likelihood of [student] replication success after an initial failure to replicate
-->
```
# Introduction

Imagine the scenario where you are a new graduate student, and you run a replication of a study in the literature that you are interested in building upon in your research.
The replication fails: perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a p-value of .3.
Or perhaps the interaction is numerically in the wrong direction, and also the main effects look different.
Alternatively the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.
Whatever the details of the replication failure, you are left with a question: Should you *try again* and run a re-replication, or should you *give up* and pick a different study to replicate and build on?

Running studies can be costly, so scientists must decide how to allocate their limited resources in the face of an uncertain literature.
Attempting to replicate existing studies is a common choice.
Scientists replicate studies for a variety of reasons, including calibrating their equipment, confirming a result before building on it, and engaging in meta-scientific practices [other reasons to highlight?
is there any citation for this?]
.
We focus on this middle motivation, where a scientist wants to build on a result, and thus wants it to succeed, but does not have strong reasons to believe or disbelieve the finding: it isn't a robust, well-tested effect, nor is it something they have particular suspicions about.
In an uncertain literature, we might not want to commit resources to building on an effect without first checking that we, with our resources and participant sample, can get the core effect we wish to build on.
Thus, scientists may want to re-replicate failed replications if the re-replication is likely to work and is indicative that future experiments in this tradition are likely to be informative.

@@

# Methods

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students have conducted replication projects, as reported in @boyce2023.
In Fall 2023, students in PSYCH 251 were offered the option to do a "rescue" project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
We report on the result of `r num_proj` rescue projects that opted to be part of the paper and completed data collection.

@@

## Sample

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that had received a subjective replication success score of 0, .25 or .5 (on a 0-1 scale) in @boyce2023, where the replication had a github repository available (github repositories were used starting in academic year 2015-2016), and where the original experiment had 200 or fewer participants (to ensure we could afford to increase the sample size).
We then contacted the replication project authors for permission to share their report and github repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects.
`r length(chosen)-num_proj` students took an incomplete or did not indicate interest in being part of the rescue paper, leaving a final sample of `r num_proj` rescue projects.

@@

## Procedure

Students conducted their rescue projects over the course of the 10-week class.
Once they had chosen a project we gave them access to the original replicators' write-up and repository, which often included the data, experiment code, and analytic code.
In many cases, students were also given the contact information of the original replicator (a few original replicators opted not to be contacted by students).

@@

With one exception, samples were collected on Prolific (the rescue of @yeshurun2003 ran in-person on the Stanford student subject pool).
We tried to power studies adequately (with a target of 2.5x original following @simonsohn2015), but due to cost constraints, not all studies were powered at this level.
The rescues had on average `r round(m,2)` times the original sample post-exclusion (median: `r round(mid,2)`, IQR: `r round(low,2)` - `r round(high,2)`, minimum: `r round(lowest, 2)`, maximum: `r round(highest,2)`, see Table \@ref(tab:tab-size) for all sample sizes).
Across the `r num_proj-1` Prolific studies, we spent \$`r pluck(cost,1,1) |> round()`, for an average of \$`r (pluck(cost,1,1)/(num_proj-1)) |> round()` per project.

## Pre-registration

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is at LINK, individual pre-registrations are linked from LINK.

@@

## Coding of results

We followed @boyce2023 in the properties of the studies we measure and how we quantify replication success.

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP.
Disagreements were resolved through discussion.
As a compliment to the subjective rating of overall success, we stastically compared one key measure of interest for each study, following @boyce2023. In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, replication, and re-replication.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We also recorded the same set of potential correlates that were used in @boyce2023 for original, replication, and rescue (these were already rated for original and replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

@@

## Overall replication rate

```{=html}
<!--* We will report the distribution of subjective replication success in our rescue sample

* If there is a mixture of projects that succeed and fail to replicate the original results, we will qualitatively describe differences that may have played a role. -->
```
All rescue projects were rated holistically for how well they replicated the original results.
We thought about replication in terms of how confident one would be to build on the line of work given the replication results, rather than focusing on any singular numeric result or significance cut-off.
All projects were rated both by the instructor (MCF) and by one of the TAs (VB or BP); the interrater reliability was `r irr |> round(3)`.

@@

Another successful replication was that of @ngo2019.
Here, the original study found a large effect.
The first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people.
It failed to find the effect.
The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separate reasons that slight changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic, and could potentially benefit from a re-replication.

@@

(TODO other successes or failures we want to discuss? ) (Ben: Emily's project was interesting: the rescue succeeded but with big stimulus-level random effects. Verity's project was also interesting: the previous replication, as well as another in the literature, showed similar but smaller findings and it really looked like she could get the effect with a bigger sample, but she ended up finding totally different results.)

@@

Small replication samples relative to original and rescue could be due to both a) powering a replication according to a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample.
Since relative sizes of the studies may play a role in replication success and how probative replications are, we show the sample sizes in Table \@ref(tab:tab-size).

@@

Overall, we do not have a clear picture of why certain studies replicated in the rescue sample and others did not, other than a few cases where fixing sample size issues may have helped.

@@

As a complement to the holistic, qualitative replication ratings used above, we also statistically compared the effect sizes of the rescue, original replication, and original study on one key measure per study.
We followed @boyce2023 in determining a key measure for each study.
When we were aware of additional direct replications (either from other class projects, or external replications in the literature), we also consider the effect sizes obtained in these additional replications.

We standardized all effect sizes into standardized mean difference (SMD) units.
One potential issue with comparisons using SMD is that noisier measures will have smaller standardized effect sizes even if the effect on the original scale is the same.
In general, the replication and rescue effect sizes were smaller than the original effect sizes, and in a few cases the effects were in the opposite direction (Figure \@ref(fig:smd)).

@@

Scientists' intuitions about whether a replication is successful and whether an effect provides support for a hypothesis (including heuristic cutoffs like p\<.05) do not always align with measures of statistical consistency [@patil2016].
For instance, two studies may both find that condition 1 results in a significantly higher outcome measure than condition 2, but the effect magnitudes may be sufficiently different that they are statistically unlikely to have come from the same population.
On the other hand, one study may find a statistically significant, but imprecisely estimated effect in a small sample, and second study may find a near-zero (null) effect, but the effect estimates from the two studies may be statistically compatible, despite one supporting a hypothesized difference and the other not.

[Ben: I think the paragraph below could use more explanation about what $p$-original is, but I'm not confident enough in my understanding of it to write it myself.]

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that two effects come from the same distribution.
As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found).
Because there is only a small number of replications per study, we impute the heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

@@

As secondary measures, we calculated the $p$-original values between a) the original and the rescue, b) the original and non-rescue replications, and c) the rescue and other replications.
For a) the rescue versus the original, median value of $p$-original was `r rescue_summ$median |> round(2)` [IQR: `r rescue_summ$low |> round(2)` - `r rescue_summ$high|> round(2)`], and `r round(rescue_summ$less_than*100)`% of the $p$-original values were less than .05.
For b) the non-rescue replications versus the original, median value of $p$-original was `r rep_summ$median |> round(2)` [IQR: `r rep_summ$low |> round(2)` - `r rep_summ$high |> round(2)`], and `r round(rep_summ$less_than*100)`% of the $p$-original values were less than .05.
For c) the rescue versus the other replications, median value of $p$-original was `r rescue_rep_summ$median |> round(2)` [IQR: `r rescue_rep_summ$low |> round(2)` - `r rescue_rep_summ$high |> round(2)`], and `r round(rescue_rep_summ$less_than*100)`% of the $p$-original values were less than .05.

@@

<!--* We will use p-original to evaluate how consistent the original effect size is with the totality of replications. We expect there to be a small number of replications, so we will impute the heterogeneity value as in https://osf.io/preprints/psyarxiv/dpyn6/. -->

Overall, allowing for a heterogeneity level of $\tau$=.21 SMD, a number of rescues and replications are not statistically consistent with the original effects.
The pattern of inconsistency does not align with which studies were rated as having replicated.
In all cases, the point estimate of the re-replication is smaller than, or in the opposite direction of, the original effect on the key measure of interest.

@@

# Discussion

We presented the results of `r num_proj` new replications that attempted to "rescue" previous failed replications reported in @boyce2023 by identifying and ameliorating possible causes of non-replication.
`r nrow(success)` of these re-replications (`r round(nrow(success)/num_proj*100)`%) mostly or fully replicated.

We don't have qualitative or quantitative explanations for why the specific studies did or did not replicate.
In some cases, increasing sample size and fixing internal validity issues in the replication seems to have led to a successful rescue (although we can't establish causality even in these cases).
However, there were other cases where the original replications had issues with a small sample or deviations in the implementation, and the re-replication addressed these issues but still failed to replicate the original results.
We can't predict what replication failures are likely to resolve given another, more thoughtful try, beyond that suggestion that glaring problems and low samples may sometimes be resolvable.

The re-replications all showed smaller effect sizes than their original studies, regardless of whether the pattern of effects replicated.
A large minority of replications had effect sizes that were statistically inconsistent with the original effect, even allowing for some heterogeneity between studies.
These diminished and inconsistent effects suggest that even if a re-replication "works", it may be difficult to build upon as follow-up studies will need large samples to detect small effects.

The reported rescue projects are a small sample of replications.
They are also chosen non-randomly, as they have been selected for twice by student interest.
However, this selection bias is likely to correlate with how graduate students choose what topics to work on and what studies to build on.
Nonetheless, with a small sample all our estimates are highly uncertain.

@@

We opened this paper with a question about what an early-career psychology researcher should do given a failed replication: should they try again or move on?
From our sample of testing the "try again" approach, it seems that the odds of a re-replication working are low (consistent with @ebersole2020).
Especially if there is not a clear, identifiable reason for the first replication's failure, another try is unlikely to rescue the original result.

# Acknowledgements {.unnumbered}

# Author Contributions {.unnumbered}

# References {.unnumbered}

<!-- Use this magic to place references here. -->

::: {#refs}
:::
